---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Load the data
```{r loadData, echo=TRUE}
unzip("activity.zip")
activity <- read.csv("activity.csv", na.strings = 'NA')
```
## What is mean total number of steps taken per day?
- Aggregate steps by date. Here is first few rows of result
```{r}
sum_by_date <- aggregate(activity$steps, by= list(activity$date), FUN = sum)
names(sum_by_date) <- c("Date","Steps")
head(sum_by_date)
```

- Histogram of the total number of steps taken each day
```{r hist, echo=T}
hist(sum_by_date$Steps, xlab = "Steps", main = "Histogram of Steps By Date")
```

- Mean of the total number of steps taken per day
```{r}
mean(sum_by_date[!is.na(sum_by_date$Steps),2])
```

- Median of the total number of steps taken per day
```{r}
median(sum_by_date$Steps,na.rm = T)
```

## What is the average daily activity pattern?
```{r}
avr <- aggregate(activity[!is.na(activity$steps), 1] , by= list(activity[ !is.na(activity$steps), 3]), FUN = mean)
names(avr) <- c("Interval", "Steps")
plot(avr, type = "l", xlab = "5-minute interval", ylab = "Average number of steps taken accross all days")
```

- Interval with maximum of steps
```{r}
avr[ avr$Steps == max(avr$Steps), ]
```
## Imputing missing values
- Total number of missing values in the dataset 
```{r}
incompelte <- activity[is.na(activity$steps), ]
length(incompelte$steps)
```

- Create a new dataset that is equal to the original dataset but with the missing data filled in with average steps of that interval.
```{r}
incompelte$steps <- avr[ avr$Interval %in% incompelte$interval, 2]
new_activity <- activity
new_activity[is.na(new_activity$steps), 1] <- avr[ avr$Interval %in% incompelte$interval, 2]
```

- Aggregate steps by date
```{r}
new_sum_by_date <- aggregate(new_activity$steps, by= list(new_activity$date), FUN = sum)
names(new_sum_by_date) <- c("Date","Steps")
new_sum_by_date
```

- Histogram of the total number of steps taken each day
```{r newhist, echo=T}
hist(new_sum_by_date$Steps, xlab = "Steps", main = "New Histogram of Steps By Date")
```

- Mean of the total number of steps taken per day
```{r}
mean(new_sum_by_date[!is.na(new_sum_by_date$Steps),2])
```

- Median of the total number of steps taken per day
```{r}
median(new_sum_by_date$Steps,na.rm = T)
```
## Are there differences in activity patterns between weekdays and weekends?

- Aggregate steps by date
```{r}
Sys.setlocale("LC_TIME", "C")
isweekend_sum_by_date <- new_activity
isweekend_sum_by_date$weekdays <- as.factor(weekdays(as.POSIXct(new_activity$date)));
isweekend_sum_by_date$isWeekend <- as.factor(ifelse(isweekend_sum_by_date$weekdays %in% c("Saturday","Sunday"), "weekend","weekday"));
isweekend_sum_by_date <- aggregate(new_activity$steps, by= list(new_activity$interval, isweekend_sum_by_date$isWeekend), FUN = mean)
names(isweekend_sum_by_date) <- c("Interval","isWeekday","Steps")

head(isweekend_sum_by_date)
```

- Histogram of the total number of steps taken each day
```{r, echo=T}
# install.packages("ggplot2")

library(ggplot2)
g <- ggplot(isweekend_sum_by_date,  aes( x = Interval, y = Steps)) 
g  + geom_line( aes(group = isWeekday) ) + facet_grid( isWeekday ~ .) + scale_x_discrete( breaks = seq(from = 0, to= max(isweekend_sum_by_date$Interval), by = 250))
```


